---
title: "how to build a color palette from any custom image"
author: "Andrea Cirillo"
date: "17/5/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# how to build a color palette from any custom image
## (and yeah, there is a package for that)

gancio con storia

[image comparing point and click website with R code]

##what you are going to find here
the post will give you a clear view of how the task was automated, providing detail of the clustering algorithm employed. It will moreover show you a small R package which was developed to ease the application of this process from other users. Finally use cases for the considered algo will be provided, from the most intuitive ggplot2 employment to the more advanced one.
## behind the scenes: k-nn for image processing
I have to be honest: my first thinking when dealing with this task was not the one I am going to show you. I was actually getting lost following a bad path mad of triples of Red Green and Blues, trying to select from frequency tables the most representative ones. But suddenly, after reading a blog post about k-nn algorithm the idea was born: let k-nn divide the picture into homogeneous clusters, having each one a "center" constituted by an RGB triple and from this way define a palett representative of the picture itself.

### the k-nn algorithm

### applying the k-nn algorithm to an image

#### read the image

#### run k-nn

### dealing with results

### show the palette

## paletter package to replicate the process
One of the best gifts the R community gave to me is the tons of packages which helped me with nearly every one of my data mining activity. This is why I often try to translate my code into packages, especially when it contains some functions which could be useful for someone. This I think is the case for the processing steps we have seen before, as we will see later on. 
For this reason you can find on Github a packge named paletter, which is currently with just one not-so-badly documented function. Paletter is freely available (of course...) and you can install it running the following>

```{r install_paletter}
library(devtools)
install_github("andreacirilloac/paletter")
```

once you will be done with the installation the actual application of the algo will be even easier, requiring you to run the *palette_maker* function, which takes as input the path to the picture and the number of colours you are wishing to derive from it:

```{r palette_maker_example}
download.file("https://andreacirilloac.github.io/dataviz/images/sacra_famiglia_canigiani.jpg", "image.jpg")
colours_vector <- palette_maker("image.jpg", number_of_colors = 32)
```


## use cases
### how to use the palette in ggplot

### possible alternative use cases
